Abstractions and exemplars: the measure noun phrase alternation in German

### 2017-12-31 Latex/sections/2descriptive.tex
Revisions in Sec.3 almost and in Sec.4 cmpltly done.

\begin{exe}
%  \ex\label{ex:intro:pseudopartitive}
%  \begin{xlist}
    \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[ein Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
    we drink a glass this wine\\
    \glt We drink a glass of this wine.}
%    \ex[*]{\label{ex:intro:pseudopartitive2} Wir trinken [[eine Tasse]\Sub{Acc} [einen leckeren Kaffee]\Sub{Acc}]\Sub{Acc}.}
%  \end{xlist}
\end{exe}

@@

% namely:
% 
% \begin{enumerate}[i.]
%   \item{\label{it:intro:idiot} alleged alternatives to the \NACa\ and the \PGCa,}
% %  \item{\label{it:intro:datsg} alternative \textit{weak} forms of dative singular neuter adjectives,}
% %  \item{\label{it:intro:femsg} case syncretism in feminine NPs,}
%   \item{\label{it:intro:plurl} similar constructions with plural/collective kind nouns,}
%   \item{\label{it:intro:noifl} grammaticalised non-inflected measure nouns, and}
%   \item{\label{it:intro:other} alternative constructions for expressing quantities.}
% \end{enumerate}
%
% \vspace{-1\baselineskip}

@@

% As for (\ref{it:intro:datsg}), a complication with neuter kind nouns in the dative is mentioned by \citet[20--22]{Zimmer2015}.
% In the \NACa, the adjective normally inflects with the so-called \textit{strong} case and number suffixes, which are used when no determiner with strong inflection is present in the NP.
% \cite{Zimmer2015} reports a high number of occurrences of adjectives being inflected with the \textit{weak} inflectional marker, which are normally only used if a strongly inflected determiner precedes the adjective: \textit{kalt-en Wasser} `cold water' instead of \textit{kalt-em Wasser}.%
% \footnote{See Section~\ref{sec:analyses} for more discussion of adjectival inflection.}
% In my corpus sample, this tendency was not nearly as clear, and there was a high number of very noisy sentences among those potentially showing this pattern.
% Hence, I do not discuss these forms.%
% \footnote{All corpus data and scripts used for this paper will be released freely, so further examination of the few cases maybe showing this kind of inflection is possible.}
% 
% \label{page:femininesyncretism}
% With feminine kind nouns, (\ref{it:intro:femsg}) is relevant.
% Whereas the singular masculine and neuter nominal subsystem is a four case system (at least for NPs with adjectives and determiners), feminine singular NPs show nominative-accusative and dative-genitive syncretisms.
% Table~\ref{tab:syncretisms} shows this as well as the partial syncretism in the plural, where no gender distinctions are made.
% This means that with dative measure nouns, the alternation is unobservable if the kind noun is feminine, and it is problematic to speak of a \textit{genitive} kind noun construction in this case.
% This is taken into account in the statistical modeling and the discussion presented in Section~\ref{sec:corpusstudies}.
% 
% \begin{table}
%   \centering
%   \begin{tabular}{llll}
%      & Masculine\slash Neuter & Feminine & Plural \\
%      \midrule
%      nominative & rot-er Wein    & \multirow{2}{*}{frisch-e Sahne}   & \multirow{2}{*}{klein-e Äpfel} \\
%      accusative & rot-en Wein    &                                   &                                \\
%      dative     & rot-em Wein    & \multirow{2}{*}{frisch-er Sahne}  & klein-en Äpfel-n               \\
%      genitive   & rot-en Wein-es &                                   & klein-er Äpfel                 \\
%   \end{tabular}
%   \caption{Case syncretisms in NPs with an adjective and without a determiner (\textit{roter Wein} `red wine', \textit{frische Sahne} `fresh cream', \textit{kleine Äpfel} `small apples')}
%   \label{tab:syncretisms}
% \end{table}

Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
%\footnote{More concretely, if these cases were included in the regression analysis of the corpus data along with a factor encoding the number of the kind noun, this factor would most assuredly override any other regressor for the data points with a plural kind noun.}
While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.

Third, some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
They cannot be used in normal MNPs.}
%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{exe}
  \ex[ ]{\label{ex:intro:pseudopartitive1}\gll Wir trinken [[ein Glas]\Sub{Acc} [dieses Weins]\Sub{Gen}]\Sub{Acc}.\\
  we drink a glass this wine\\
  \glt We drink a glass of this wine.}
\end{exe}

@@

@@

Second, we find that, if the kind noun is a plural count noun as in \textit{ein Sack kleine Äpfel} (\NACa) or \textit{ein Sack kleiner Äpfel} (\PGCa) `a bag of small apples', a similar alternation between PGC and NAC can be observed.
In line with experimental results reported in \citet[15--16]{Zimmer2015}, I found that the PGC is so dominant with plural kind nouns (794 of 861 cases, or over 92\%, cf.\ Section~\ref{sec:corpusstudies}) that the alternation cannot be analysed in the same way as in the singular.
While this will play a role in the interpretation of the corpus findings, MNPs with plural kind nouns will not be included in the corpus study and the experiments reported in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.

Third, some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions such as \textit{Pfunde verlieren} `lose pounds' and \textit{Prozente machen} `make a profit'.
They cannot be used in normal MNPs.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.

@@

### 2017-12-31 Latex/sections/3corpus.tex
Revisions in Sec.3 almost and in Sec.4 cmpltly done.

\begin{table}
  \centering
  \resizebox{\textwidth}{!}{
    \begin{tabular}{llrlrrrc}
    Model level  & Regressor         & $\text{p}_{\text{PB}}$ & Factor level & Coefficient & CI low & CI high & CI excludes 0 \\
    \midrule
    First        & Badness           &                 &              &       &  &   &              \\
                 & Cardinal          &                 & No           &       &  &   &              \\
                 & Genitives         &                 &              &       &  &   &              \\

    Second       & (Kindfreq)        &                 &              &       &  &   &              \\
    (Kind)       & (Kindgender)      &                 &              &       &  &   &              \\[0.5\baselineskip]

    Second       & Measureclass      &                 & Container    &       &  &   &              \\
    (Measure)    &                   &                 & Rest         &       &  &   &              \\
                 &                   &                 & Amount       &       &  &   &              \\
                 &                   &                 & Portion      &       &  &   &              \\
                 & (Measurefreq)     &                 &              &       &  &   &              \\

  \end{tabular}
  }
  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is XXX}
  \label{tab:smalltable}
\end{table}

The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as describing a context typical of the \PGCd.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
Regressors which did not reach sig=0.05 in in the pbkrtest (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.

@@

Then, a multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
It was estimated at -4.328.

The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as \textit{group level predictors} (or \textit{second-level effects}), cf.\ \citet[265--269,302--304]{GelmanHill2006} and Section~\ref{sec:itemandexemplareffects}.
The same goes for those listed with the kind lemma as their unit of reference.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

\begin{table}
  \centering
  \resizebox{\textwidth}{!}{
    \begin{tabular}{llrlrrrc}
    Model level  & Regressor         & $\text{p}_{\text{PB}}$ & Factor level & Coefficient & CI low & CI high & CI excludes 0 \\
    \midrule
    First        & Badness           &  0.002                 &              & -0.152      & -0.247 & -0.061  & *             \\
                 & Cardinal          &  0.001                 & No           &  1.189      &  0.862 &  1.466  & *             \\
                 & Genitives         &  0.001                 &              & -0.693      & -0.768 & -0.592  & *             \\
                 & Measurecase       &  0.001                 & Acc          &  0.030      & -0.150 &  0.212  &               \\
                 &                   &                        & Dat          &  0.705      &  0.455 &  0.944  & *             \\[0.5\baselineskip]

    Second       & Kindattraction    &  0.020                 &              &  0.225      &  0.049 &  0.393  & *             \\
    (Kind)       & Kindfreq          &  0.095                 &              &  0.146      & -0.023 &  0.301  &               \\
                 & Kindgender        &  0.001                 & Neut         &  0.021      & -0.367 &  0.392  &               \\
                 &                   &                        & Fem          &  1.269      &  0.800 &  1.709  & *             \\[0.5\baselineskip]

    Second       & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
    (Measure)    & Measureclass      &  0.001                 & Container    &  0.252      & -0.265 &  0.788  &               \\
                 &                   &                        & Rest         &  0.421      & -0.209 &  1.063  &               \\
                 &                   &                        & Amount       &  0.831      &  0.215 &  1.432  & *             \\
                 &                   &                        & Portion      &  1.217      &  0.675 &  1.684  & *             \\
                 & Measurefreq       &  0.005                 &              & -0.231      & -0.363 & -0.079  & *             \\

  \end{tabular}
  }
  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -4.328}
  \label{tab:bigtable}
\end{table}

Standard diagnostics show that the model quality is quite good.%
\footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
I am convinced that the model should be specified based on theoretical considerations (and known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
Removal of useless regressors (judging by, for example, the pbkrtest and coefficients of determination) is the only type of model selection I allow in my research.}
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and none of the corrected $\text{GVIF}^{1/2\text{df}}$ was higher than 1.6.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
Only \textit{Kindfreq} ($\mpPB=0.095$) can be seen as slightly too high to be convincing, failing at sig=0.05.

Using collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) was not successful.
While the attraction measures reach satisfying sig levels, the $\text{p}_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
The coefficients of determination drop to $R_m^2=0.383$ and $R^2_c=0.490$.

@@

%\hl{As collostructional analysis is relatively popular in some cognitively oriented areas of corpus linguistics, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
%The corresponding regressor did not even reach sig=0.1 in the pbkrtest.
%Both the cognitive underpinnings and the math behind collostructional analysis are still much debated (see prominently \citealp{SchmidKuechenhoff2013,KuechenhoffSchmid2015}), I see this result }

>>>>

\begin{table}
  \centering
  \resizebox{\textwidth}{!}{
    \begin{tabular}{llrlrrrc}
    Model level  & Regressor         & $\text{p}_{\text{PB}}$ & Factor level & Coefficient & CI low & CI high & CI excludes 0 \\
    \midrule
    1                & Badness           & 0.042                  &              &  0.120      & -0.002 &  0.228  &              \\
                     & Cardinal          & 0.001                  & No           &  1.419      &  0.869 &  1.993  &  *           \\
                     & Genitives         & 0.001                  &              & -0.710      & -0.815 & -0.556  &  *           \\[0.5\baselineskip]

    2 (Kindlemma)    & (Kindfreq)        & (0.781)                &              &             &        &         &              \\
                     & (Kindgender)      & (0.199)                &              &             &        &         &              \\[0.5\baselineskip]

    2 (Measurelemma) & Measureclass      & 0.005                  & Container    & 0.445       & -0.579 &  1.532  &              \\
                     &                   &                        & Rest         & 1.745       &  0.649 &  2.831  &  *           \\
                     &                   &                        & Amount       & 1.597       &  0.125 &  2.751  &  *           \\
                     &                   &                        & Portion      & 1.782       &  0.771 &  2.690  &  *           \\
                     & (Measurefreq)     & (0.265)                &              &             &        &         &              \\

  \end{tabular}
  }
  \caption{Coefficient table with 95\% bootstrap confidence intervals for the pre-study; the intercept is -5.370}
  \label{tab:smalltable}
\end{table}

The coefficient estimates are specified in Table~\ref{tab:smalltable} for each regressor (or regressor level) in the columns labelled \textit{Coefficient}.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as characterising a configuration typical of the \PGCd.
For a robust quantification of the precision of the estimation, I ran a parametric bootstrap (using the \mbox{\textit{confint.merMod}} function from \textit{lme4}) with 1,000 replications and using the percentile method for the calculation of the intervals.
The resulting 95\% bootstrap confidence intervals are reported in Table~\ref{tab:smalltable} in the columns labelled \textit{CI low} and \textit{CI high} (= upper and lower 2.5th percentiles).
The column \textit{CI excludes 0} shows an asterisk for those intervals that do not include 0.
Furthermore, for each regressor, a p-value was obtained by dropping the regressor from the full model, re-estimating the nested model, and comparing it to the full model.
Instead of inexact Wald approximations and Likelihood Ratio Tests, I used a drop-in bootstrap replacement for the Likelihood Ratio Test from the function \textit{PBmodcomp} from the \textit{pbkrtest} package \citep{HalekohHojsgaard2014}.
I call the corresponding value $p_{\text{PB}}$, and it is given in the respective columns in Table~\ref{tab:smalltable}.
Regressors which did not reach sig=0.05 in in the PBmodcomp test (\textit{Kindfreq}, \textit{Kindgender}, and \textit{Measurefreq}) were removed from the model, appear in brackets in Table~\ref{tab:smalltable}, and have no coefficient estimates.
The overall intercept is -5.370 and reoresents \textit{Cardinal=Yes}, \textit{Measureclass=Physical} and 0 for all (z-transformed) numeric regressors.
The standard deviation in the random intercepts is 1.157 for \textit{Measurelemma} ($p_{\text{PB}}=0.001$) and 0.923 for \textit{Kindlemma} ($p_{\text{PB}}=0.0.002$).
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.278$ and $R^2_c=0.566$ (see \citealp{Gries2015} for a basic introduction to these $R^2$ measures, or else \citealp{NakagawaSchielzeth2013}).
This means that the effect of the fixed effects (marginal $R^2$) is weak to intermediate, and that lemma-specific effects (conditional minus marginal $R^2$) are equally strong with a difference between the two coefficients of 0.289.

This show that the main prototypical factors \textit{Cardinal} and \textit{Measureclass} indeed play the predicted role
The abstract non-referential physical measure nouns (such as \textit{Gramm} `gram' or \textit{Liter} `litre') with a high degree of grammaticalisation favour the \NACb.
At the other end of the scale, nouns denoting natural portions like \textit{Haufen} `heap', \textit{Bündel} `bundle', \textit{Schluck} `gulp' favour the \PGCd.
Also, The presence of a cardinal modifier clearly favours the \NACb.
The stylistic factor \textit{Genitives} is confirmed to have an influence as predicted inasmuch as a general lack of genitives (which is encoded as a \textit{higher} value of the variable) favours the \NACb.
\textit{Badness} points into the wrong direction, but it barely reaches sig=0.05, and its confidence interval includes 0.
The lemma frequency regressors both clearly fail the PBmodcomp test.
I will return to these results in Section~\ref{sec:interpretation}.

@@

Then, a multilevel logistic regression model was fit which models the influence of the regressors specified in Table~\ref{tab:variables} on the probability that the \PGCa\ is chosen over the \NACa.
All regressors from Table~\ref{tab:variables} were included, and the measure lemma and the kind noun lemma were specified as varying-intercept random effects.
The sample size was \textit{n=}5,063 with 1,134 cases of \PGCa\ and 3,929 cases of \NACa.
The results of the estimation are shown in Table~\ref{tab:bigtable} and in Figure~\ref{fig:fixeffs}.
The intercept comprises \textit{Cardinal=Yes}, \textit{Measurecase=Nom}, \textit{Kindgender=Masc}, \textit{Measureclass=Physical}, and 0 for all numeric z-transformed regressors.
It was estimated at -3.548.

The regressors with the measure lemma as their unit of reference have no within-measure lemma variance, and the \textit{glmer} function automatically estimates them as group-level predictors (or second-level effects), cf.\ \citet[265--269,302--304]{GelmanHill2006} and Section~\ref{sec:itemandexemplareffects}.
The same goes for those listed with the kind lemma as their unit of reference.
Given the coding of the response variable, coefficients leaning to the positive side can be interpreted as favouring the \PGCa.

\begin{table}
  \centering
  \resizebox{\textwidth}{!}{
    \begin{tabular}{llrlrrrc}
    Model level  & Regressor         & $\text{p}_{\text{PB}}$ & Factor level & Coefficient & CI low & CI high & CI excludes 0 \\
    \midrule
    1               & Badness           &  0.002                 &              & -0.152      & -0.247 & -0.061  & *             \\
                    & Cardinal          &  0.001                 & No           &  1.189      &  0.862 &  1.466  & *             \\
                    & Genitives         &  0.001                 &              & -0.693      & -0.768 & -0.592  & *             \\
                    & Measurecase       &  0.001                 & Acc          &  0.030      & -0.150 &  0.212  &               \\
                    &                   &                        & Dat          &  0.705      &  0.455 &  0.944  & *             \\[0.5\baselineskip]

    2 (Kindlemma)   & Kindattraction    &  0.020                 &              &  0.225      &  0.049 &  0.393  & *             \\
                    & Kindfreq          &  0.095                 &              &  0.146      & -0.023 &  0.301  &               \\
                    & Kindgender        &  0.001                 & Neut         &  0.021      & -0.367 &  0.392  &               \\
                    &                   &                        & Fem          &  1.269      &  0.800 &  1.709  & *             \\[0.5\baselineskip]

    2 (Measurelemma & Measureattraction &  0.001                 &              &  0.282      &  0.106 &  0.447  & *             \\
                    & Measureclass      &  0.001                 & Container    &  0.252      & -0.265 &  0.788  &               \\
                    &                   &                        & Rest         &  0.421      & -0.209 &  1.063  &               \\
                    &                   &                        & Amount       &  0.831      &  0.215 &  1.432  & *             \\
                    &                   &                        & Portion      &  1.217      &  0.675 &  1.684  & *             \\
                    & Measurefreq       &  0.005                 &              & -0.231      & -0.363 & -0.079  & *             \\

  \end{tabular}
  }
  \caption{Coefficient table with 95\% bootstrap confidence intervals for the main study; the intercept is -3.548}
  \label{tab:bigtable}
\end{table}

Standard diagnostics show that the model quality is quite good.%
\footnote{Notice that no aggressive model selection was applied, especially not upward model selection including probing for interactions.
I am convinced that the model should be specified based on theoretical considerations (and known nuisance effects) to avoid the dangers of data dredging and to avoid models which are difficult to interpret.
Removal of useless regressors (judging by, for example, the PBmodcomp test and coefficients of determination) is the only type of model selection I allow in my research.}
Nakagawa \& Schielzeth's pseudo-coefficient of determination is $R_m^2=0.409$ and $R^2_c=0.495$.
The rate of correct predictions is 0.843, which means a proportional reduction of error of $\lambda=0.297$.
Generalised variance inflation factors for the regressors were calculated to check for multicollinearity \citep{FoxMonette1992,ZuurEa2010}, and the highest corrected $\text{GVIF}^{1/2\text{df}}$ was 1.520 for \textit{Cardinal}.
The lemma intercepts have standard deviations of $\sigma_{\text{Measurelemma}}=0.448$ and $\sigma_{\text{Kindlemma}}=0.604$.
Only \textit{Kindfreq} ($\mpPB=0.095$) could be seen as slightly too high to be convincing, failing at sig=0.05.

Using signed logarithmised collexeme strength (\textit{Measurecollo} and \textit{Kindcollo}) (with smoothing to avoid log(0) operations) instead of the quotient for the attraction strength (\textit{Measureattraction} and \textit{Kindattraction}) (see Section~\ref{sec:variablesandannotation}) was not successful.
While the attraction measures reach satisfying sig levels in the PBmodcomp test (0.020 for \textit{Kindattraction} and 0.001 for \textit{Measureattraction}), the $p_{\text{PB}}$ value for \textit{Kindcollo} was 0.191 and the one for \textit{Measurecollo} was 0.443.
The coefficients of determination drop to $R_m^2=0.135$ and $R^2_c=0.636$.

@@

%\hl{As collostructional analysis is relatively popular in some cognitively oriented areas of corpus linguistics, I tried using collexeme strength as a regressor (with smoothing to remedy the mathematical problems), and the results were unsatisfactory compared to the simple quotient used here.
%The corresponding regressor did not even reach sig=0.1 in the PBmodcomp test.
%Both the cognitive underpinnings and the math behind collostructional analysis are still much debated (see prominently \citealp{SchmidKuechenhoff2013,KuechenhoffSchmid2015}), I see this result }
